Stability and Chaos in Multi Agent Reinforcement Learning

### 2020-10-06 aamas2021-formatting-instructions/aamas21-formatting-instructions/sample.tex
Updated Discussion in AAMAS Submission

@@

@@

\acmSubmissionID{???}

@@

\title[InstabilityinMARL]{Instability in Multi Agent Reinforcement Learning}

@@

This study allows for parameters to be appropriately chosen to ensure the safe convergence of a learning algorithm and as a first step towards understanding the range of behaviours that can be displayed by learning using the Q-Learning algorithm. We validate our theoretical results through numerical simulation and show that, within the bounds of experimental error, the region of instability can be characterised by the learning dynamics. We also note a few important implications of the study: 1) convergence is rare in Multi-Agent Reinforcement Learning. It is found that convergence to a unique fixed point occurs for simple games with 2 players and 2 actions. However, once either of these increase, convergence no longer occurs. 2) the existence of instability is not affected by the structure of the game itself. In fact, at least in the competitive regime, the payoff structure does not affect whether the game converges or not. 3) Convergence is guaranteed for low step-lengths but rapidly decreases as the step length increases. 
\end{abstract}

@@

\keywords{Auction Theory, Reinforcement Learning, Social Simulation}

@@

\paragraph{Contribution}
In this study, we will be considering the question: under what choice of parameters is a
multi-agent system, which is trained on an iterated game using Q-Learning, likely to converge to an
equilibrium as opposed to displaying complex, or chaotic behaviours? In investigating this question, we will make the
following assumptions:

@@

Analyses have been performed from the point of view of the replicator dynamic towards the various types of behaviours that can emerge from iterated games. Such games are considered imperfect, in that agents do not make decisions based on a rational process derived from perfect information of the game, but rather through attempting to anticipate their opponents' behaviour based on experience \cite{Galla2011}. \cite{Imhof2005} presents the observation that, in the game of Prisoner's Dilemma, cyclic behaviour emerges in which the optimal strategy cycles across defection, cooperation and tit-for-tat despite defection being the NE. \cite{Galla2011} builds on this notion by showing that, in fact, such behaviour is highly dependent on the presence of memory loss in the system. In the absence of memory loss, the learning converges to the NE as expected. However, as this parameter is increased, the stability of the NE is removed and other stable points are created.

Taking this notion even further, \cite{Sato2002} presents a study in which it is shown that learning, even on a simple two player game (they choose rock-paper-scissors), demonstrates Hamiltonian chaos (see \cite{Strogatz2000} for details on chaos). In this regime, there are no stable equilibria and the game will never converge to an equilibrium. To analyse this phenomena further, \cite{Galla2013} describes an investigation of two-person iterated games in which the players learn using \textit{Experience Weighted Attraction} (EWA), a form of reinforcement learning whose analysis typically falls under the purview of experimental economics \cite{Camerer2009}. Through numerical simulations, the authors were able to show that, once again, the emergence of chaos and cycles, is dependent on the choice of parameters. A rigorous theoretical analysis, which follows the techniques presented in \cite{Opper1992}, of the replicator dynamic corresponding to EWA presented a method for characterising the regions in parameter space in which a game is likely to converge, exhibit chaos or limit cycles. More recently, the authors presented an extended study in \cite{Sanders2018}, in which this analysis was extended towards a generic $p$-player game, which showed that chaotic dynamics are more likely to be observed as the number of players \textit{p} increases, whilst a similar study \cite{Pangallo2019} indicates that an increase in the number of actions $N$ in the strategy space similarly increases the likelihood of unstable behaviour.

@@

\paragraph{Dynamics of RL} The earliest example of such an analysis may be found in \cite{Borgers1997}, in which the \textit{Cross Learning} algorithm is analysed and it is shown that, in a continuous time limit, the learning model converges to the replicator dynamic. Adapting this work for the benefit of the AI community, the authors of \cite{Tuyls2006AnGames} similarly derive a relation between Q-Learning (see \cite{Barber2012}) and the replicator dynamic. In doing so, the authors present an evolutionary model which accurately predicts the dynamics of Q-Learning. This sprung forward a vast array of literature which perform similar derivations for various reinforcement learning algorithms. \cite{Bloembergen2015} presents an excellent review of these for the interested reader. In particular it should be noted that many of these studies focus on normal form games, in which the payoff matrix does not change (i.e. the environment is static). Recent work, such as \cite{Hennes2008}, extends this work towards multi-state games, though this is restricted to 2 state games. Similarly, \cite{Galstyan2013}, extends the work in \cite{Tuyls2006AnGames} towards games with continuous action spaces. Most recently, \cite{Hu2019} considers Q-learning in the mean-field limit, in which the strategies of populations of agents are considered.

The analysis of RL from an EGT perspective is a growing field, with a significant scope for expansion, as is discussed in our conclusion. 
However, the relation between RL and the replicator dynamic allows for the assumption of convergence in reinforcement learning to be lifted and instead for the emergence of cyclic behaviour and chaos to be examined. In particular, this study aims to perform a similar analysis to that in \cite{Sanders2018} and establish the regions in parameter space in which Q-Learning converges to stable fixed points and where the learning is unstable. To the best of our knowledge, no such work exists in this context. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

    As discussed in the previous section, when learning on iterated games, player
    strategies may exhibit much more complex behaviour than
    convergence to a Nash Equilibrium (NE), \st{These behaviours include:}
    including convergence to a unique
    equilibrium (though not always to an NE), convergence to one of
    multiple equilibria, limit cycles and chaos. These behaviours are
    illustrated in Fig.\ref{fig::DynamicalBehaviours}. To be able to predict the behaviour of a learning algorithm, it should ideally converge to a stable equilibrium
    although it is still possible to study systems with multiple
    equilibria or limit cycles \cite{Strogatz2000}. However, it would
    be difficult to control systems whose dynamics are governed by
    chaos (though research into controlling chaos is ongoing and rife
    with opportunity \cite{Fradkov2009}). It would, therefore, be a useful endeavour to
    determine the conditions under which these sorts of behaviours
    arise.

@@

Here, $\alpha$ and $\tau$ are the parameters of the agent as described at the end of this section. Agent 1 (resp.~agent 2) takes action $i$ with probability
$x_i$ (resp.~$y_j$).
\st{while Agent 2 takes action $j$ with probability $y_j$}. When these actions are taken, the agents
receive payoff $a_{ij}$ and $b_{ji}$ respectively. With these equations, it is possible to
predict the expected behaviour of Q-Learning agentswhich they go on to verify.

@@

Our first point of call is to extend the dynamics of Tuyls et al.~in  (\ref{eqn::EOM}a) and (\ref{eqn::EOM}b) \st{and extend it} to a general $p$-player game. This yields the dynamics (\ref{eqn::pEOM}). Here, player $\mu$ chooses action $i_{\mu}$ from its action space $\mathcal{A}$ at time $t$ with probability $\xmu{i}{\mu}$ and receives a reward $\payoff{i}{\mu}$ from its payoff matrix $P^\mu$ depending on its own action and the actions of all other agents $i_{-\mu}$. Formally, the notation $i_{-\mu}$ denotes the set $\{ i_\kappa : \kappa \in \{1, 2, ..., p\} \backslash \{\mu\} \}$ and $-i_{\mu}$ denotes the set $\mathcal{A} \backslash {i_\mu}$.

@@

\begin{equation}
	\begin{split}
	Z(\Vec{\psi}) = \int D[\Vec{x}, \Vec{\hat{x}}] exp( i \sum_{i, \mu} \int dt [ \hxmu{i} (\frac{\dxmu{i}}{\xmu{i}{\mu}} - \talpha \rho_i^\mu (t) - h_i^\mu (t))]) \times \\ exp(-i \alpha \ttau \sum_{\mu} \sum_{i_\mu, i_{-\mu}} \int dt [\hxmu{i} \payoff{i}{\mu} \prod_{\kappa \neq \mu} \xmu{i}{\kappa} )])) 
    \times \\ exp(-i \alpha \ttau \sum_{\mu} \sum_{j_\mu, i_\mu, i_{-\mu}} \int dt [\hxmu{j}  \xmu{i}{\mu} \payoff{i}{\mu} \prod_{\kappa \neq \mu} \xmu{i}{\kappa}])) 
	\times \\ exp(i \sum_{i, \mu}
	\int dt[\xmu{i} \psi^\mu_i(t)]),
\end{split}
\end{equation}

The last two exponentials contain the payoffs of the game. These are randomly generated using a multivariate gaussian and then held fixed for the rest of the game. As such these comprise the aforementioned 'quenched disorder'. We will average over this quenched disorder (and therefore average over all possible realisations of payoff matrices).

@@

in which we have assumed that all players' actions are independent and drawn from the same initial distribution (\ref{eqn::Payoffs})
and therefore dropped the distinction between players and strategy components. The terms $G, C, \eta_1, \eta_0$ are correlation functions, generated during when averaging the Gaussian. These are given by

@@

It is important to note the effect of the former assumption (that all actions of all agents are i.i.d) is substantial. With this assumption, we reduce the second summation term, which couples actions from all agents $\xmu{i}{\mu}$ in (\ref{eqn::EOM}) to the uncoupled term $\sqrt{2} \alpha \htau \eta_0(t)$ in (\ref{eqn::EffectiveDynamics}). This is a strong assumption which is required to analyse the stability of the dynamics in the competitive regime. However, as shown by the experimental evaluation, it does not produce a strong discrepancy in describing the qualitative effect on stability caused by the parameters $\alpha, \Gamma, \tau$.

@@

where $\chi = \int dt' G(t - t')$ (at a fixed point where the probabilities do not vary with time, this is equal to $\int dt' G(t, t')$), $\eta_0(t) = q^{n/2}z$ and $z$ is drawn from a Gaussian of zero mean and unit variance and $n = p - 1$. Having arrived at a condition for the fixed point, we then examine the long-term behaviour of the perturbation $\xpert$. To do this, we take the Fourier transform of (\ref{eqn::Linearised}) and analyse its behaviour at $\omega = 0$. After some manipulation, which are given in the Supplementary Material, this gives

@@

\begin{enumerate}
    \item Convergence is rare. This is seen by the fact that, for all allowed choices of $\Gamma, \alpha, \tau$, the right hand side of (\ref{eqn::Final}) is positive, which violates the stability criterion. This implies that games, in general, will not converge.
    \item The probability of convergence decreases with increasing $\alpha$, regardless of the choice of $N, p$. We can see this since the right hand side of (\ref{eqn::Final}) tends further away from zero (i.e. further from the region of stability).
    \item At low values of $\alpha$, the game is everywhere convergent, regardless of the choice of $\Gamma$. In fact, the choice of $\Gamma$ does not affect the outcome of the stability analysis at all. This is more easily interpreted from the heatmapts in Figure \ref{fig:theory}.
\end{enumerate}

@@

We illustrate these implications in Figure \ref{fig:theory} which plots the value (and takes the absolute value for convenience) obtained by the right hand side of (\ref{eqn::Final}). In order to determine these values, we first had to solve (\ref{eqn::fixed_point}) for \st{the values of} $q$ and $\chi$. We ignore the choice $\xfixed  = 0$, since the proportion of agents who choose a non zero action probability is considered to be 1 \fb{the meaning of this is not clear to me}, an intuitive result \fb{assumption?} which is also considered in \cite{Sanders2018} and \cite{Coolen2005}. As such, we solve for $\xfixed$ using a Newton-Raphson root-finding approach and use the given \fb{obtained} expressions for $q, \chi$ to determine their value \fb{can we give such expressions?}. It should note that this method of finding $\xfixed$ yields only an approximate value and so the $q, \chi$ which are found are estimates of their true value.

We evaluate (\ref{eqn::Final}) for varying values of $\alpha$ and $\Gamma$, whilst keeping $\tau$ fixed at 0.05 and restricting the range of $\Gamma$ to $[-1, 0]$, as it is only for this range that the solution to (\ref{eqn::fixed_point}) is defined. The range of $\alpha$ is also restricted to $[0.01, 0.05]$ as it is found that for larger values \st{of $\alpha$,} convergence is not seen.

\fb{Equation} (\ref{eqn::Final}) always evaluates to a positive value, suggesting that Q-learning in games is always unstable. We believe that this is due to the heavy assumptions made during the derivation of the analytic result, the strongest of which was to drop the discrepancy between players and actions, treating each action of each player to be equal \fb{in which sense equal?}. However, even within the error that this assumption generates, it is still clear (and is verified experimentally) that convergence in Q-Learning is rare. We therefore interpret (\ref{eqn::Final}), for given parameters, by determining how close the right hand side is to the stability boundary (i.e., how close the value is to zero). A choice of parameters which evaluates close to the stability boundary has a higher chance of being convergent than one which lies further from the boundary. Therefore, the heat-maps in Figure \ref{fig:theory} should be considered to show the probability of convergence for a choice of $(\alpha, \Gamma)$. Darker regions denote a higher probability of convergence whereas lighter regions are those whose values are further from zero and therefore show a lower probability of convergence.

In addition, we can visualise the implications 2 and 3 in Sec.~\ref{sec::Theory} by comparing the heat-maps. We see in all of the heat-maps that the probability of convergence decreases as $\alpha$ increases (i.e., moving from left to right). At low values of $\alpha$, convergence is seen for all choices of $\Gamma$. However, as $\alpha$ increases, the overall likelihood of convergence rapidly decreases and 
%the convergence 
it is more sensitive to the choice of $\Gamma$. A point to note is the implication that, with the increase of $p$ and $N$ comes a significant increase in the likelihood of convergence. This implication experimental evaluation \fb{?} and therefore can be seen as a limitation of (\ref{eqn::Final}). Instead, it should be viewed that, for a choice of $p, N$, the variation in stability of a system can be determined from $\alpha$ and $\Gamma$.

@@

To verify \fb{experimentally the theoretical results}, and to examine the underlying structure of stability and chaos in
Multi-Agent Q-Learning, we perform a series of numerical experiments by varying the parameters $\Gamma$ and
$\alpha$ whilst keeping $\tau$ fixed. The purpose 
\st{of this experiment} is to determine a heuristic estimate for the Lyapunov exponent, which measures the degree to which initial conditions which start close to one another converge or diverge as the dynamics evolve \cite{Strogatz2000}. The former indicates a stable system (in the sense of Lyapunov \cite{}) whilst the latter signals instability. The procedure for calculating the Lyapunov exponent is given in (\ref{eq::LiapExpo}).

To generate the numerical simulations in Figure~\ref{fig:NumericalExperiments} we used the
following procedure.
\begin{enumerate}
   \item Fix the paremeters $\tau, \gamma$. The latter is held at 0.5 in all experiments as (\ref{eqn::EOM}) indicates that the value of $\gamma$ does not affect the overall behaviour of the system.
   \item Initialise values of $\Gamma, \alpha$. These will be swept over in the experiment.
\item Generate 5 payoff matrices for both agents by sampling from a multi-variate Gaussian 
(variables are the payoff elements) with mean zero and covariance parameterised by $\Gamma$.
\item Initialise 2 sets of agents (denote their action probabilities as $\Vec{A_1}$ and $\Vec{A_2}$ respectively) each with random initial conditions (i.e., random action probabilities).
\item Allow both sets of agents to learn over a maximum of $1 \times 10^4$ iterations.
\item After 500 iterations measure the distance $\Vec{\delta_1} = |A_2 - A_1|$
\item After 10000 iterations measure the distance $\Vec{\delta_n} = |A_2 - A_1|$
\item Determine \fb{the Lyapunov exponent} 
\begin{eqnarray*}
\label{eq::LiapExpo}
    \lambda_n & = & \max_{i} 10^{-4} ln(\delta_{n, i}/\delta_{1, i})
\end{eqnarray*}
for each action $i$ and average this value over all five realisations of the payoff matrix. 
\end{enumerate}

This yields the result shown in Figure~\ref{fig:NumericalExperiments}. We see from this that the stability of the system is highly dependent on
the value of $\alpha$ and (although less so) on $\Gamma$.

>>>>

\usepackage{soul}
\usepackage{subcaption}

@@

\usepackage{xcolor} % REMOVE THIS AT THE END
\newcommand\fb[1]{\textcolor{blue}{FB: #1}}

@@

\acmSubmissionID{489}

@@

\title[InstabilityinMARL]{Stability and Chaos in Multi Agent Reinforcement Learning}

@@

This study allows for parameters to be appropriately chosen to ensure the safe convergence of a learning algorithm and as a first step towards understanding the range of behaviours that can be displayed by learning using the Q-Learning algorithm. We validate our theoretical results through numerical simulation and show that, within the bounds of experimental error, the region of instability can be characterised by the learning dynamics.
\end{abstract}

@@

\keywords{Game Theory, Reinforcement Learning, Dynamical Systems}

@@

\paragraph{Contribution}
In this study, we will be considering the question: under what choice of parameters is a
multi-agent system, which is trained on an iterated game using Q-Learning, likely to converge to an
equilibrium as opposed to displaying complex, or unstable behaviours? In investigating this question, we will make the
following assumptions:

@@

Analyses have been performed from the point of view of the replicator dynamic towards the various types of behaviours that can emerge from iterated games. Such games are considered imperfect, in that agents make decisions by attempting to anticipate their opponents' behaviour based on experience \cite{Galla2011}. \cite{Imhof2005} presents the observation that, in the game of Prisoner's Dilemma, cyclic behaviour emerges in which the optimal strategy cycles across defection, cooperation and tit-for-tat despite defection being the NE. \cite{Galla2011} builds on this notion by showing that, in fact, such behaviour is highly dependent on the presence of memory loss in the system. In the absence of memory loss, the learning converges to the NE as expected. However, as this parameter is increased, the stability of the NE is removed and other stable points are created.

Taking this notion even further,  \cite{Galla2013} describes an investigation of two-person iterated games in which the players learn using \textit{Experience Weighted Attraction} (EWA), a form of reinforcement learning whose analysis typically falls under the purview of experimental economics \cite{Camerer2009}. Through numerical simulations, the authors were able to show that, once again, the emergence of chaos and cycles (see \cite{Strogatz2000} for details on chaotic and periodic motion), is dependent on the choice of parameters. A rigorous theoretical analysis, which applies the techniques presented in \cite{Opper1992}, of the replicator dynamic corresponding to EWA presented a method for characterising the regions in parameter space in which a game is likely to converge, exhibit chaos or limit cycles. More recently, the authors presented an extended study in \cite{Sanders2018}, in which this analysis was extended towards a generic $p$-player game, which showed that chaotic dynamics are more likely to be observed as the number of players \textit{p} increases.

@@

\paragraph{Dynamics of RL} In \cite{Tuyls2006AnGames}, the authors are able to derive a relation between Q-Learning (see \cite{Barber2012}) and the replicator dynamic. In doing so, the authors present an evolutionary model which accurately predicts the dynamics of Q-Learning. This sprung forward a vast array of literature which perform similar derivations for various reinforcement learning algorithms. \cite{Bloembergen2015} presents an excellent review of these. In particular it should be noted that many of these studies focus on normal form games, in which the payoff matrix does not change (i.e. the environment is static). Recent work, such as \cite{Hennes2008}, extends this work towards multi-state games, though this is restricted to 2 state games. Similarly, \cite{Galstyan2013}, extends the work in \cite{Tuyls2006AnGames} towards games with continuous action spaces. Most recently, \cite{Hu2019} considers Q-learning in the mean-field limit, in which the strategies of populations of agents are considered.

The relation between RL and the replicator dynamic allows for the assumption of convergence in reinforcement learning to be lifted and instead for the emergence of more complex behaviour to be examined. In particular, this study aims to perform a similar analysis to that in \cite{Sanders2018} and establish the regions in parameter space in which Q-Learning converges to stable fixed points and where the learning is unstable. To the best of our knowledge, no such work exists in this context. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

    As discussed in the previous section, when learning on iterated games, player
    strategies may exhibit much more complex behaviour than
    convergence to a Nash Equilibrium (NE),
    including convergence to a unique
    equilibrium (though not always to an NE), convergence to one of
    multiple equilibria, limit cycles and chaos. These behaviours are
    illustrated in Fig.\ref{fig::DynamicalBehaviours}. To be able to predict the behaviour of a learning algorithm, it should ideally converge to a stable equilibrium
    although it is still possible to study systems with multiple
    equilibria or limit cycles \cite{Strogatz2000}. However, it would
    be difficult to control systems whose dynamics are governed by
    chaos (though research into controlling chaos is ongoing and rife
    with opportunity \cite{Fradkov2009}). It would, therefore, be a useful endeavour to
    determine the conditions under which these sorts of behaviours
    arise.

@@

Here, $\alpha$ and $\tau$ are the parameters of the agent as described at the end of this section. Agent 1 (resp.~agent 2) takes action $i$ with probability
$x_i$ (resp.~$y_j$).
When these actions are taken, the agents
receive payoff $a_{ij}$ and $b_{ji}$ respectively. With these equations, it is possible to
predict the expected behaviour of Q-Learning agents which they go on to verify.

@@

Our first point of call is to extend the dynamics of Tuyls et al.~in  (\ref{eqn::EOM}a) and (\ref{eqn::EOM}b) to a general $p$-player game. This yields the dynamics (\ref{eqn::pEOM}). Here, player $\mu$ chooses action $i_{\mu}$ from its action space $\mathcal{A}$ at time $t$ with probability $\xmu{i}{\mu}$ and receives a reward $\payoff{i}{\mu}$ from its payoff matrix $P^\mu$ depending on its own action and the actions of all other agents $i_{-\mu}$. Formally, the notation $i_{-\mu}$ denotes the set $\{ i_\kappa : \kappa \in \{1, 2, ..., p\} \backslash \{\mu\} \}$ and $-i_{\mu}$ denotes the set $\mathcal{A} \backslash {i_\mu}$.

@@

\begin{equation}
\label{eqn::separatePayoffs}
	\begin{split}
	Z(\Vec{\psi}) = \int D[\Vec{x}, \Vec{\hat{x}}] exp( i \sum_{i, \mu} \int dt [ \hxmu{i} (\frac{\dxmu{i}}{\xmu{i}{\mu}} - \talpha \rho_i^\mu (t) - h_i^\mu (t))]) \times \\ exp(-i \alpha \ttau \sum_{\mu} \sum_{i_\mu, i_{-\mu}} \int dt [\hxmu{i} \payoff{i}{\mu} \prod_{\kappa \neq \mu} \xmu{i}{\kappa} )])) 
    \times \\ exp(-i \alpha \ttau \sum_{\mu} \sum_{j_\mu, i_\mu, i_{-\mu}} \int dt [\hxmu{j}  \xmu{i}{\mu} \payoff{i}{\mu} \prod_{\kappa \neq \mu} \xmu{i}{\kappa}])) 
	\times \\ exp(i \sum_{i, \mu}
	\int dt[\xmu{i} \psi^\mu_i(t)]),
\end{split}
\end{equation}

The last two exponentials contain the payoffs of the game. These are randomly generated using a multivariate gaussian and then held fixed for the rest of the game. As such these comprise the aforementioned 'quenched disorder'. We will average over this quenched disorder (and therefore average over all possible realisations of payoff matrices). We define $Q$ to be the product of the second and third exponentials in (\ref{eqn::separatePayoffs})

\begin{equation}
\begin{split}
        <Q> = exp(- \frac{\alpha^2 \ttau ^2}{2} N \sum_{\mu} \Big [ \int dt dt' L^\mu(t, t') \prod_{\kappa \neq \mu} C^\kappa (t, t') + \\ \Gamma \sum_{\nu \neq \mu} K^\mu (t, t') K^\nu (t, t') \prod_{\kappa \not\in \{\mu, \nu\}} C^\kappa (t, t') \Big ] ) \times \\
        exp(- \frac{\alpha^2 \htau ^2}{2} N \sum_{\mu} \Big [ \int dt dt' L^\mu(t, t') C^\mu (t, t') \prod_{\kappa \neq \mu} C^\kappa (t, t') + \\ \Gamma \sum_{\nu \neq \mu} A^{\mu \nu} (t, t') C^\nu (t, t') C^\mu (t, t') \prod_{\kappa \not\in \{\mu, \nu\}} C^\kappa (t, t') \Big ] )
\end{split}
\end{equation}

In which

\begin{align}
\label{eqn::correlations}
    \begin{split}
        C^\mu (t, t') & :=  N^{-1} \sum_i \xmu{i}{\mu} \xmudash{i}{\mu} \\
        L^\mu (t, t') & :=  N^{-1} \sum_i \hxmu{i} \hxmudash{i} \\
        K^\mu (t, t') & :=  N^{-1} \sum_i \xmu{i}{\mu} \hxmudash{i} \\
        A^{\mu, \nu} (t, t') & :=  N^{-1} \sum_i \hxmu{i} \hxnudash{i}.
    \end{split}
\end{align}

@@

in which we have assumed that all players' actions are independent and drawn from the same initial distribution (i.i.d)
and therefore dropped the distinction between players and strategy components. The terms $G, C, \eta_1, \eta_0$ are correlation functions, generated during when averaging the Gaussian. These are given by

@@

It is important to note the effect of the former assumption (that all actions of all agents are i.i.d) is substantial. With this assumption, we identify the coupled term $A^{\mu, \nu} (t, t')$ in (\ref{eqn::correlations}) with the uncoupled term $L^\mu (t, t')$ (since there is no distinction between the actions of agents $\mu$ and $\nu$). This is a strong assumption which removes the interdependency between agents in the analysis. However, as shown by the experimental evaluation, it does not produce a strong discrepancy in describing the qualitative effect on stability caused by the parameters $\alpha, \Gamma, \tau$.

@@

where $\chi = \int dt' G(t - t')$, $\eta_0(t) = q^{n/2}z$ and $z$ is drawn from a Gaussian of zero mean and unit variance and $n = p - 1$. This gives us an expression for calculating $\xfixed$. We disregard the choice $\xfixed  = 0$, since we assume that no action will be chosen with exactly zero probability, an intuitive assumption which is also considered in \cite{Sanders2018} and \cite{Coolen2005}. The expression inside the squared bracket admits a positive solution only in the region $\Gamma \in [-1, 0]$, whilst for positive $\Gamma$, the nature of solutions may not be guaranteed. Therefore, we restrict our analysis to this region.

Having arrived at a condition for the fixed point, we then examine the long-term behaviour of the perturbation $\xpert$. To do this, we take the Fourier transform of (\ref{eqn::Linearised}) and analyse its behaviour at $\omega = 0$. After some manipulation, which are given in the Supplementary Material, this gives

@@

\begin{enumerate}
    \item A trivial result is that the game is everywhere convergent (i.e. regardless of the choice of $\Gamma, N, p$) for the cases of $\alpha$ = 0 and/or $\tau$ = 0.  We see that choosing these values would result in the right hand side of (\ref{eqn::Final}) yielding a value of zero which falls within the stable region.
    \item Convergence is rare. This is seen by the fact that the analytic result overestimates the region of instability. In fact, for all allowed choices of $\Gamma, \alpha, \tau$, the right hand side of (\ref{eqn::Final}) is negative, which violates the stability criterion. This implies that games, in general, will not converge.
    \item The likelihood of convergence decreases with increasing $\alpha$ and $\tau$, regardless of the choice of $N, p$. We can see this since the right hand side of (\ref{eqn::Final}) tends further away from zero (i.e. further from the region of stability).
    \item The choice of $\Gamma$ does not affect the stability of the game in the negatively-correlated regime. Therefore, the only dependent factors are $\alpha, \tau, N, p$. This is more easily interpreted from the heatmaps in Figure \ref{fig:theory}.
    \item The likelihood of convergence decreases as $p$ increases. We see this by noticing the weighting that the $(p-1)$ in the second term of (\ref{eqn::Final}) places. As $p$ increases, this term drives the system further from the stability boundary.

\end{enumerate}

@@

We illustrate these implications in Figure \ref{fig:theory} which plots the value obtained by the right hand side of (\ref{eqn::Final}). In order to determine these values, we first had to solve (\ref{eqn::fixed_point}) for $q$ and $\chi$. We determine the $\xfixed$ which solves the expression inside the square brackets of (\ref{eqn::fixed_point}) using a Newton-Raphson root-finding approach and use the following expressions for $q, \chi$ to determine their value.

\begin{equation*}
    \begin{split}
        1 = \frac{1}{\sqrt{2 \pi}} \int_{-\infty}^{\infty} \xfixed(z) exp(-\frac{z^2}{2}) dz \\ 
        q = \frac{1}{\sqrt{2 \pi}} \int_{-\infty}^{\infty} \xfixed(z)^2 exp(-\frac{z^2}{2}) dz \\ 
        \chi = \frac{1}{q^{n/2}}\frac{1}{\sqrt{2 \pi}} \int_{-\infty}^{\infty} \frac{\partial \xfixed (z)}{\partial z} exp(-\frac{z^2}{2}) dz
    \end{split}
\end{equation*}

It should note that this method of finding $\xfixed$ yields only an approximate value and so the $q, \chi$ which are found are estimates of their true value.

We consider each of the implications 2-5 from the previous section in turn, choosing to forego illustrating 1 as it can be seen trivially from the expression. In those heatmaps for which $\tau$ is held fixed \textbf{denote them}, we choose $\tau = 0.05 $ and vary $\Gamma$ in the range $[-1, 0]$ and $\alpha$ in the range $[0.01, 0.05]$. For those in which $\alpha$ is held fixed, \textbf{denote them} we choose $\alpha =  0.05$ and vary $\tau$ in the range $[0.01, 0.05]$.

By looking at the values of the heatmap, we see that, for all choices of the parameters, other than $\alpha = 0, \tau = 0$, (\ref{eqn::Final}) evaluates to a negative value, which signals instability. As such, the analysis over-estimates the region of instability. We believe that this is due to the heavy assumptions made during the derivation of the analytic result, the strongest of which was to drop the discrepancy between players and actions, treating each action of each player to be independent and identically distributed. However, even within the error that this assumption generates, it is still clear (and is verified experimentally) that convergence in Q-Learning is rare. Indeed as the choice of $N$ increases, the stability boundary determined by experiment (Figure \ref{fig:NumericalExperiments}) tends towards $\alpha = 0, \tau = 0$.

Similarly, as the value of $N$ increases, the values in the heatmap (seen on the right of the map) reduce. This is due to the assumption made in the derivation of the analytic result that $N \rightarrow \infty$. As the value of N increases, the discrepancy between the true behaviour of the system and analytic result decreases. This does, however, give rise to a limitation of the result - namely that the effect of the number of actions $N$ cannot be accurately inferred from the (\ref{eqn::Final}). However, it should still be noted that, even with the increase of N, the values in parameter space remain negative and still, accurately, predict instability. We go on to verify that the effect of $p, \alpha, \Gamma$ and $\tau$ are accurately captured by (\ref{eqn::Final}).

We interpret (\ref{eqn::Final}), for given parameters, by determining how close the right hand side is to the stability boundary (i.e., how close the value is to zero). A choice of parameters which evaluates close to the stability boundary has a higher chance (by the analytic error) of being convergent than one which lies further from the boundary. Therefore, the heat-maps in Figure \ref{fig:theory} should be considered to show the probability of convergence for a choice of $(\alpha, \Gamma)$. Lighter regions denote a higher probability of convergence whereas darker regions are those whose values are further from zero and therefore show a lower probability of convergence.

We can visualise the implication 3 by noticing that the likelihood of convergence decreases (the heatmap becomes darker) as $\alpha$ increases, and similarly for $\tau$. This occurs independent of the choice of $N$ or $p$. The implication 4 can be inferred by the trend of convergence, which occurs only in varying $\alpha$ and $\tau$. The choice of $\Gamma$
makes almost no difference in the evaluation of (\ref{eqn::Final}). We anticipate that this is due to the restriction of $\Gamma$ to be in the range $[-1, 0]$. In this range, the scaling effect of $\Gamma^2$ is dominated by the $\alpha^4$ and $\tau^4$ terms which appear in (\ref{eqn::Final}).

Finally we can see implication 5, by comparing the heatmaps for different values of $p$ \textbf{denote them}. We see that, though the overall trend is the same, the heatmaps tend to be darker for higher $p$, suggesting that these lie further from the stability boundary.

@@

To verify experimentally the theoretical results, and to examine the underlying structure of stability and chaos in
Multi-Agent Q-Learning, we perform a series of numerical experiments by varying the parameters $\Gamma$ and
$\alpha$ whilst keeping $\tau$ fixed. The aim is to determine the regions in which games learnt using Q-Learning converge to an equilibrium. The results of these experiments are shown in Figure \ref{fig:NumericalExperiments}, with parameters chosen to match those in the analytic assessment (Figure \ref{fig:theory}).
% By examination it is found that, in many cases the system oscillates about some equilibrium point in a low dimensional chaotic manner (i.e. the manner in which the action probabilities change is indistinguishable from random motion) \textbf{get some graphs for this}

% The purpose 
% \st{of this experiment} is to determine a heuristic estimate for the Lyapunov exponent, which measures the degree to which initial conditions which start close to one another converge or diverge as the dynamics evolve \cite{Strogatz2000}. The former indicates a stable system (in the sense of Lyapunov \cite{}) whilst the latter signals instability. The procedure for calculating the Lyapunov exponent is given in (\ref{eq::LiapExpo}).

To generate the numerical simulations in Figure~\ref{fig:NumericalExperiments} we used the
following procedure. Note in the following that for any heatmap in which $\tau$ is swept over, the roles of $\alpha$ and $\tau$ are swapped.
\begin{enumerate}
   \item Fix the parameters $\tau, \gamma$. The former is held at 0.05 and the latter at 0.1. For the heatmaps in which $\tau$ is swept over $\alpha$ is held fixed at 0.05.
   \item Initialise values of $\Gamma, \alpha$ (or $\tau$ appropriately). These will be swept over in the experiment.
\item Generate 15 payoff matrices for both agents by sampling from a multi-variate Gaussian 
(variables are the payoff elements) with mean zero and covariance parameterised by $\Gamma$.
\item For each of these payoff matrices, initialise a set of agents with random initial conditions (i.e., random action probabilities).
\item Allow both sets of agents to learn over a maximum of $1.5 \times 10^4$ iterations.
\item Keep track of the action probabilities over a window of 500 iterations. At the end of each window, determine the percentage difference between the maximum and minimum values of each strategy component
\item If the difference is less than 1 \% consider the game converged. Otherwise continue to the next window.
\item If the game reaches $1.5 \times 10^4$ iterations without satisfying the relative distance criterion, consider it to be non-convergent. Determine the fraction of these 15 games which have converged.

% \item After 500 iterations measure the distance $\Vec{\delta_1} = |A_2 - A_1|$
% \item After 10000 iterations measure the distance $\Vec{\delta_n} = |A_2 - A_1|$
% \item Determine \fb{the Lyapunov exponent} 
% \begin{eqnarray*}
% \label{eq::LiapExpo}
%     \lambda_n & = & \max_{i} 10^{-4} ln(\delta_{n, i}/\delta_{1, i})
% \end{eqnarray*}
% for each action $i$ and average this value over all five realisations of the payoff matrix. 
\end{enumerate}

We see in Figure \ref{fig:NumericalExperiments} that the stability of the system is highly dependent on
the value of $\alpha$ and $\tau$ but not on $\Gamma$.
